Equations of state
\begin{itemize}
    \item Bag model \cite[eq. 7.33]{lecture_notes}
    \item Assumptions on speed of sound and constant parameters
    \item Giese's paper \cite{giese_model-independent_2020}
    \item Enqvist, old paper \cite{enqvist_nucleation_1992}
\end{itemize}

Employment?
- "things that need to be done"
-> suitable for public dissemination
- write some short description,
    - ArXiv paper

2 papers by Giese, second one is more complete
Paper 1 \cite{giese_2020}
Paper 2 \cite{giese_2021}

TODO: understand these
\begin{itemize}
    \item Relativistic hydrodynamics from the review article (aka. lecture notes) \cite{lecture_notes}
    \item Reviews from other people (not just Hindmarsh)
    \item Article by Mazumdar \& White, especially section 7: relativistic combustion \cite{mazumdar_review_2019}
    \item The sound shell model paper
    \item Really understand what's going on! Compute with pen \& paper etc.
    \item Relativistic hydrodynamics
    \begin{itemize}
        \item The book by Luciano Rezzolla \& Olindo Zanotti \cite{rezzolla_relativistic_2013}
        \item Read the part I (some 300 pages at first)
        \item Contains more than necessary for the thesis
    \end{itemize}
    \item The phase transition section should have plenty of content before Christmas
    \item Enqvist's paper
    \begin{itemize}
        \item Good background: connects the equations of state to particle physics
        \item -> What kinds of equations of state would a particle theory produce
        \item Assumes ultrarelativisticity, which is no longer necessary due to the improvement of computational tools
    \end{itemize}
    \item Ultimately the equation of state can be derived from the "Thermal effective potential".
    \item Chemical potential is zero, as there is no conserved particle number.
    \item J-function, fB, lecture eq. 2.19 etc.
    \begin{itemize}
        \item These will be our first equation of state.
    \end{itemize}
\end{itemize}

"My job is to give the GWs the right energy-momentum tensor"

\section{Basics}
Electroweak symmetry breaking at $10^{-11} s$, $100 -- 1000 GeV$.
First-order phase transition
-> bubbles just below the critical temperature
-> gravitational waves

Standard model has a crossover, but many extensions first-order.

Matter-antimatter asymmetry -> net baryon number
(But isn't B an accidental symmetry, whereas B-L is a fundamental one?
How does this correspond to the neutrino content of the universe?)
\cite{lecture_notes}

The gravitational wave spectrum can be calculated from a few thermodynamic properties of ultrarelativistic matter, which can be computed from quantum field theory.
These parameters can be measured by LISA.

\section{Relativistic hydrodynamics}
Ultrarelativistic particles
Non-constant particle number

Statistical mechanics \cite{huang_statistical_1987}
\cite{schroeder_introduction_2000}
\cite[ch. 4]{lecture_notes}

Densities:
Energy [density]
\begin{equation}
e = T \frac{\partial p}{\partial T} - p
\end{equation}

Enthalpy density
\begin{align}
w
&= T \frac{\partial p}{\partial T} \\
&= e+p \\
&= Ts
\end{align}

Entropy density
\begin{equation}
s = \frac{dp}{dT}
\end{equation}

Entropy current
\begin{equation}
S^\mu = su^\mu
\end{equation}
\cite[p. 23]{lecture_notes}

Energy-momentum tensor
\cites[eq. 5.23]{lecture_notes}[eq. 3]{giese_2020}
\begin{align}
T^{\mu \nu}_f
&= (e+p) u^\mu u^\nu + p g^{\mu \nu} \\
&= w u^\mu u^\nu + p g^{\mu \nu}
\end{align}

\section{Relativistic combustion}
The conservation of energy-momentum results in the bubble wall junction conditions
\begin{align}
w_- \tilde{\gamma}_-^2 \tilde{v}_- &= w_+ \tilde{\gamma}_+^2 \tilde{v}_+, \\
w_- \tilde{\gamma}_+^2 \tilde{v}_-^2 + p_- &= w_+ \tilde{\gamma}_+^2 \tilde{v}_+^2 + p_+
\end{align}
These can be rearranged as
\cites[eq. 7.32]{lecture_notes}[eq. 6-7]{giese_2020}
\begin{align}
\tilde{v_+} \tilde{v_-} = \frac{p_+ - p_-}{e_+ - e_-} \\
\frac{\tilde{v}_+}{\tilde{v}_-} = \frac{e_- + p_+}{e_+ + p_-}
\end{align}
And simplified to
\begin{align}
\frac{v_+}{v_-} &= \frac{1 - \frac{\Delta e}{w_+}}{1 - \frac{\Delta p}{w+}} \\
v_+ v_- &= \frac{\frac{\Delta p}{w_+}}{\frac{\Delta e}{w_+}}
\end{align}

Speed of sound
\cite[eq. 13]{giese_2020}
\begin{equation}
c_s^2 = \frac{dp}{de}
\end{equation}

Trace anomaly
\begin{equation}
\theta = \frac{1}{4}(e-3p)
\end{equation}

Difference of trace anomaly
\begin{align}
\Delta \theta
&= \theta_+ - \theta_- \\
&= \theta_s - \theta_b
\end{align}

Transition strength
\begin{equation}
\alpha_+ = \frac{4 \Delta \theta}{3 w_+}
\end{equation}

Enthalpy ratio
\begin{equation}
r = \frac{w_+}{w_-}
\end{equation}


\section{The bag model}
The bag model is the simplest [TODO: REF] model for the equation of state.

In general, the speed of sound is temperature-dependent.
However, if both phases are assumed to be ultrarelativistic, the speed of sound simplifies to $c_s`2 = 1/3$.
This is satisfied by the "bag" model, where
\begin{align}
p_s &= a_s T^4 - V_s \\
p_b &= a_b T^4,
\end{align}
where $a_s$, $a_b$ and $V_s$ are positive constants with $a_s > a_b$.
s=symmetric, b=broken.

\section{Types of solutions}
Subsonic deflagrations

Detonations

Supersonic deflagrations (hybrids)

\section{Energy redistribution}
Efficiency factor $\kappa$, which is the fraction of energy that is converted into fluid motion.
\begin{equation}
\kappa = \frac{3}{\chi_w^3 \Delta \theta} \int d\chi \chi^2 w \gamma^2 v^2
\end{equation}
From solving the hydrodynamic equations of state of a single expanding bubble.

Kinetic energy fraction
\cites[eq. 7.36]{lecture_notes}[eq. 5]{giese_2020}
\begin{align}
K
&= \frac{\rho_\text{fl}}{\bar{e}}
&= \frac{1}{\mathcal{v} \bar{e}} \int d^3 x w \gamma^2 v^2,
\end{align}
where $\mathcal{v}$ is the volume we average over, and $\bar{e}$ is the mean energy density.
It should be noted that $0 < K < 1$.
\cite{giese_2020}

For a single bubble we have, using its volume and the energy density of the symmetric phase that the bubble replaces, the kinetic energy fraction
\cites[eq. 7.37]{lecture_notes}[eq. 5]{giese_2020}
\begin{equation}
K_1 = \frac{3}{\chi_w^3 e_s} \int d\chi \chi^2 w \gamma^2 v^2.
\end{equation}

\section{More complex models}
The bag model includes severe approximations, and for realistic simulations a more complex model is needed.
The primary parameters for these models are
\begin{itemize}
    \item The phase transition parameters: percolation temperature, phase transition duration
    \item Bubble wall velocity
    \item Fraction of energy that is converted into fluid motion: $\kappa$
    \item Numerical prefactor from lattice simulations
\end{itemize}

Speed of sound may not be ultrarelativistic, especially in the broken phase.


\section{Extensions of the standard model}
Extensions of the Standard model \cite{caprini_detecting_2020}
(mentioned in \cite[p. 14]{lecture_notes})

\section{Noise background}
Non-thermal phenomena: GW production by preheating, when the inflaton decays to SM particles.
\cite{lecture_notes}
